---
title: "kate_analysis"
author: "Kate Colvin"
date: "2026-03-20"
output: pdf_document
---

Analysis of a Phase III Randomized Trial of SARS-CoV-2 Vaccine Efficacy and Safety Over Time

- add additional figure maybe 
- format survival results table

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(readxl)
library(survival)
library(ggsurvfit)
library(gtsummary)
library(gt)

```

Loading in data

```{r}

q2c <- read_excel("data/Q2c.xlsx", 
                    col_types = c("numeric",
                                  "numeric",
                                  "numeric",
                                  "numeric",  # needed to fix infection_time loading as TRUE/FALSE
                                  "numeric")) %>% 
  janitor::clean_names()

```


Summary Table (Q2C Table 1)
```{r}

q2c_table <- q2c %>%
  mutate(infection_factor = factor(infection, levels = c(0,1), 
                                   labels = c("Not Infected", "Infected")))

table1 <- q2c_table %>% 
  tbl_summary(include = c(enrollment_time, infection_factor, infection_time, last_fu_time), 
              by = infection_factor, 
              label = list(
                enrollment_time ~ "Time from Enrollment to Dose 2 (days)", 
                infection_time ~ "Time to Infection (days)",
                last_fu_time ~ "Last Follow-up Time (days)")) %>% 
  add_overall() %>% 
  bold_labels() %>% 
  modify_caption("Dose 2 Follow-Up and Observation Times by Infection Status")

table1

q2c_table1 <- table1 %>% as_gt()
gtsave(q2c_table1, "results/q2c_table1.png")

```

Graph of Times maybe? 

# 1. Please use the provided data set (i.e., Q2c.xlsx) to obtain the probability of getting infected within 12 months as well as the corresponding 95% confidence interval.

```{r}

# infection_time is time they were infected, if they were infected
# last_fu_time is last follow-up for those who were not infected

# Creating consistent time variable 
q2c <- q2c %>% mutate(time = ifelse(q2c$infection == 1, q2c$infection_time, q2c$last_fu_time))

# KM curve
km_fit <- survfit(Surv(time, infection) ~ 1, 
                  data = q2c)

km_year <- summary(km_fit, times = 365)
km_year

# Probability of infection within one year!
1 - km_year$surv  

```

```{r, eval = FALSE}

# Kaplan-Meier Curve, probability of survival (no infection)

survfit2(Surv(time, infection) ~ 1, 
         data = q2c) |> 
  ggsurvfit(color = "blue") +
  labs(
    x = "Days",
    y = "Overall survival probability") + 
  add_confidence_interval(fill = "blue")

```


```{r}

# Probability of infection

plot(km_fit$time, 1 - km_fit$surv, 
  type = "s", 
  col = "blue",
  xlab = "Days",
  ylab = "Cumulative Infection Probability",
  ylim = c(0, 0.5))

# Confidence intervals + line at year mark
lines(sf$time, 1 - km_fit$upper, col = "blue", lty = 2)  # upper CI
lines(sf$time, 1 - km_fit$lower, col = "blue", lty = 2)  # lower CI
abline(v = 365, col = "red", lty = 2)  # dashed red line
grid()

```


# 2. If someone asks you: “Based on the above data, what is the median time to infection, and the mean of the time to infection after the second shot?” how will you respond?

Median time - The time by which 50% of people are infected.
- NA, less than 50% of participants became infected during follow-up. 

Mean time to infection - The average infection-free time during the first 12 months is 362 days. 

```{r}

# Median
summary(km_fit)$table["median"]

# Mean
print(km_fit, rmean = 365)

```

